---
title: "Untitled"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,warning = FALSE)
```

```{r}
library(readxl)
CovidData <- read.csv('tidycovid19.csv')
library(lubridate)
CovidData$date = as.Date(parse_date_time(CovidData$date,orders=c("y","ym","ymd")))
```  

```{r}
library(rio)
Countries <- import("Countries.xlsx", setclass = "tibble")
newdata <- merge(x=CovidData,y=Countries,all.x = TRUE)
```

```{r}
library(dplyr)
Task1Data <- {newdata %>% 
    filter(continent=='Europe'|continent=='North America'|continent=='South America') %>%
    select(country,date, confirmed, gcmr_retail_recreation,   gcmr_grocery_pharmacy, continent) }
```

```{r}
firstdiff <- function(x) {
  shifted <- c(0,x[1:(length(x)-1)])
  result = x-shifted
  which_negative = which(result<0)
  result[which_negative] = NA
  return(result)
}
Task1Data <- Task1Data %>% mutate(daily_confirmed = firstdiff(confirmed))
```

```{r}
mean_ <- function(...) mean(..., na.rm=T)
max_ <- function(...) max(..., na.rm=T)
Task2Data = {Task1Data %>% group_by(continent) %>% summarise(date=date,confirmed = confirmed,gcmr_retail_recreation = gcmr_retail_recreation, gcmr_grocery_pharmacy = gcmr_grocery_pharmacy,continent=continent)}
```

```{r  fig.height=5,fig.width=9}
library(ggplot2)
retaildate<-{Task2Data %>% ggplot(aes(x=date,y=gcmr_retail_recreation,color=continent))}+geom_point(aes(fill=continent),stat = 'identity')+xlab('Date')+ylab('Activity in retail and recreation (% change from baseline)')
retaildate
```
A sharp decline must correspond to strict government restrictions

The speed of recovery seems similar (same gradient)
```{r  fig.height=5,fig.width=9}
library(ggplot2)
grophadate<-{Task2Data %>% ggplot(aes(x=date,y=gcmr_grocery_pharmacy,color=continent))}+geom_point(aes(fill=continent),stat = 'identity')+xlab('Date')+ylab('Activity in Grocery and Phamacy (% change from baseline)')
grophadate
```
general decrease was not as much as decrease in activity in retail_recreation, come back rapidly even increase positively.


```{r}
library(rio)
library(dplyr)
intuserdata<-import('internet_users.xlsx',setclass='tibble')
```
```{r}
library(tidyr)
Tidy_Data <- pivot_longer(intuserdata, !country, names_to= 'year', values_to = 'percentage')
tidydata<-filter(Tidy_Data,year %in% c(1980:2018))
```
```{r}
newuserdata <- merge(x=tidydata,y=Countries,all.x = TRUE)
newuserdata<- filter(newuserdata,continent=='Europe'|continent=='North America'|continent=='South America')
```
```{r fig.height=5,fig.width=9}
mean_ <- function(...) mean(..., na.rm=T)
averageline<-{newuserdata %>% group_by(continent)%>% summarise(mean_percentage=mean_(percentage),year=year)}

```
```{r fig.height=5,fig.width=9}
library(ggplot2)
intuserplot<-ggplot(newuserdata,aes(x=year,y=percentage,group=continent,color=continent,shape=continent))+geom_point()+labs(x='Year',y='percentage of internet user') 

print(intuserplot)
```

same pattern:  Europe is most common of internet user. followed by north america(a wider range).South America has relatively concentrated distribution of internet user.

Want to show change of activity in retail and grocery v.s percentage of internet users (must be more ecommerce implies less risky to be confirmed)
How to link with activity in retail and grocery? because the time line is different.

