---
title: "Google Trends Planning Report"
author: Max Wilkinson
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

I am going to be using Google Trends [1] data on the search volume for various terms over time. Google Trends is an unbiased sample of Google search data. It’s anonymised, categorized and aggregated. I will be looking at the interest over time of various search terms to indicate the effect lockdowns and government interventions have had on consumer behavior. 

The first search terms I will be investigating are `restaurants` and `take-out`. Below is a graph of the average interest for these terms throughout European, and North, Central and South American countries. The y axis represents search interest relative to the highest point on the chart for the given region and time. A value of 100 is the peak popularity for the term. A value of 50 means that the term is half as popular.

```{r chunk1}
library(dplyr)
library(readxl)
library(lubridate)
RestaurantTrendData <- read.csv("gt_restaurants.csv")
TakeOutTrendData <- read.csv("gt_take_out.csv")

RestaurantTrendData <- RestaurantTrendData %>%
  mutate(date = as.Date(parse_date_time(RestaurantTrendData$Week,c("d/m/y"))))
TakeOutTrendData <- TakeOutTrendData %>%
  mutate(date = as.Date(parse_date_time(TakeOutTrendData$Week,c("d/m/y"))))

library(ggplot2)
plot_restuarants_trends <- ggplot(RestaurantTrendData, aes(x = date, y = interest)) +
  geom_line(aes(color='red')) +
  geom_line(data=TakeOutTrendData, aes(color='blue', labels='restaurants')) +
  xlab("Date") +
  ylab("Interest") +
  ggtitle("Interest in Restaurants and Take-Out have a inverse relationship", 
          subtitle="Y-axis represents search interest relative to the highest point") +
  scale_x_date(date_breaks = "months", date_labels = "%m-%y") + 
  scale_color_manual(labels = c("take-out", "restaurants"), values = c("blue", "red")) +
  guides(color=guide_legend("Search Term"))


print(plot_restuarants_trends)
```
The plot above shows that the two search terms' interest have an inverse relationship. As we can see, when lockdown and restrictions began to be imposed by governments, consumer interest in `restaurants` decreased whereas interest in `take-out` increased at a similar magnitude. This shows that in lockdowns, consumers are looking to buy more food online, due to sit-in restaurants being closed. This indicates that we, OmniCorp, should focus on online retail and hospitality when lockdown measures are introduced. It is also important to note that the interest in `restaurants` increased back to normal levels almost as fast as it fell. However, it had a smaller, more steady decrease as a second wave of restrictions took place.

The next graph shows interest in the search term `restaurants` for the top 7 countries for GDP in Europe, and North, Central and South America. This selection may be appropriate as it could be said that countries with a greater GDP will have more expendable income and thus spend more in the retail and hospitality sectors.

```{r chunk2}
RestaurantCountriesTrendData <- read.csv("gt_restaurants_countries.csv")

RestaurantCountriesTrendData <- RestaurantCountriesTrendData %>%
  mutate(date = as.Date(parse_date_time(RestaurantCountriesTrendData$Week,c("d/m/y"))))
RestaurantCountriesTrendData <- subset(RestaurantCountriesTrendData, select = -c(Week))

library(reshape2)
RestaurantCountriesTrendDataReshaped <- melt(RestaurantCountriesTrendData, "date")

plot_restuarants_countries_trends <- ggplot(RestaurantCountriesTrendDataReshaped, aes(x = date, y = value, color=variable)) +
  geom_line() +
  xlab("Date") +
  ylab("Interest") +
  ggtitle("Interest in Restaurants for top GDP countries follow similar trend", 
          subtitle="Y-axis represents search interest relative to the highest point") +
  scale_x_date(date_breaks = "months", date_labels = "%m-%y") +
  #scale_color_manual(labels = c("take-out", "restaurants"), values = c("blue", "red")) +
  guides(color=guide_legend("Country"))


print(plot_restuarants_countries_trends)
```
As we can see from the graph above, when interest for `restaurants` is split into individual countries, it follows a similar pattern as for countries as a whole. There are two countries in particular that had a low point in interest for a relatively long time compared to the other countries. These are the United Kingdom and Italy. They both stayed in a prolonged trough of interest for roughly 6 weeks compared to the other countries that rebounded much quicker. However, both these countries had interest going back to pre-lockdown figures around the end of August. The only country for this not to happen is Brazil. One suggestion for this may be due to longer and more strict restrictions in this country.

# Reference

[1] Google Trends. (2015). Google Trends. [online] Available at: https://trends.google.com/trends/.
