---
title: "Max Draft Report"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Using [Google Trends](https://trends.google.com/trends/) data on the search volume for various terms, we can investigate the general interest over time. Google Trends is an unbiased sample of Google search data. It’s anonymised, categorized and aggregated. For the regions we are looking at - Europe, North, Central and South America - the [percentage of population](https://www.internetworldstats.com/) that uses the internet is 88%, 95%, 61% and 72% respectively. This indicates that the search patterns shown by Google Trends may be an accurate representation of the behaviors and interests of these regions.

We will investigate the search terms `restaurants` and `take-out`, the two can be seen as the compliment of one another. We will plot a graph of the average interest for these terms throughout European, North, Central and South American countries. The y axis represents search interest relative to the highest point on the chart for the given region and time. A value of 100 is the peak popularity for the term. A value of 50 means that the term is half as popular.

```{r chunk1}
library(dplyr)
library(readxl)
library(lubridate)
RestaurantTrendData <- read.csv("gt_restaurants.csv")
TakeOutTrendData <- read.csv("gt_take_out.csv")

RestaurantTrendData <- RestaurantTrendData %>%
  mutate(date = as.Date(parse_date_time(RestaurantTrendData$Week,c("d/m/y"))))
TakeOutTrendData <- TakeOutTrendData %>%
  mutate(date = as.Date(parse_date_time(TakeOutTrendData$Week,c("d/m/y"))))

library(ggplot2)
plot_restuarants_trends <- ggplot(RestaurantTrendData, aes(x = date, y = interest)) +
  geom_line(aes(color='red')) +
  geom_line(data=TakeOutTrendData, aes(color='blue', labels='restaurants')) +
  xlab("Date") +
  ylab("Interest") +
  ggtitle("Interest in Restaurants and Take-Out have a inverse relationship", 
          subtitle="Y-axis represents search interest relative to the highest point") +
  scale_x_date(date_breaks = "months", date_labels = "%m-%y") + 
  scale_color_manual(labels = c("take-out", "restaurants"), values = c("blue", "red")) +
  guides(color=guide_legend("Search Term"))


print(plot_restuarants_trends)
```
The plot above shows that the two search terms' interest have an inverse relationship. As we can see, when lockdown and restrictions began to be imposed by governments, consumer interest in `restaurants` decreased whereas interest in `take-out` increased at a similar magnitude. This shows that in lockdowns, consumers are looking to buy more food online, possibly due to sit-in restaurants being closed or people scared to go out in public. This indicates that we, OmniCorp, should focus on online retail and hospitality when lockdown measures are introduced. It is also important to note that the interest in `restaurants` increased back to normal levels almost as fast as it fell. However, it had a smaller, more steady decrease as a second wave of restrictions took place whereas `takeout` did not have a similar magnitude of increase.